---
title: "Individual fluidity of self-reported sexual identity in Stockholm County from 2010 to 2021"
author: Guoqiang Zhang
email: guoqiang.zhang@ki.se
output: html_notebook
editor_options: 
  chunk_output_type: console
---

Notes:

This R Notebook illustrates changes in sexual identity from 2010 to 2021 among participants in the pooled cohort across SPHC 2002, 2006, and 2010, who provided data on sexual identity in all three follow-up survey years (2010, 2014, and 2021).

### 1. Load Packages
```{r}
library(haven)
library(tidyverse)
library(naniar)
library(ggalluvial)
library(ggbreak)
library(ggrepel)
```

### 2. The Stockholm Public Health Cohort
#### 2.1. SPHC 2002
```{r}
d_2002 <- read_sas("/Volumes/LGBT Project data/sphc02_07_10_14_21.sas7bdat")

# sexual identity in 2010
table( d_2002$F10F103, useNA = "always" )
d_2002$F10F103[ d_2002$F10F103 == 9 ] <- NA
d_2002$sexual_identity_2010 <- factor( ifelse( d_2002$F10F103 == 1, "Heterosexual",
                                                  ifelse( d_2002$F10F103 == 2, "Homosexual",
                                                          ifelse( d_2002$F10F103 == 3, "Bisexual", "Other" ) ) ),
                                          levels = c( "Heterosexual", "Homosexual", "Bisexual", "Other" ) )
table( d_2002$sexual_identity_2010, useNA = "always" )

# sexual identity in 2014
table( d_2002$F14F103, useNA = "always" )
d_2002$sexual_identity_2014 <- factor( ifelse( d_2002$F14F103 == 1, "Heterosexual",
                                                  ifelse( d_2002$F14F103 == 2, "Homosexual",
                                                          ifelse( d_2002$F14F103 == 3, "Bisexual", "Other" ) ) ),
                                          levels = c( "Heterosexual", "Homosexual", "Bisexual", "Other" ) )
table( d_2002$sexual_identity_2014, useNA = "always" )

# sexual identity in 2021
table( d_2002$F21F91, useNA = "always" )
d_2002$sexual_identity_2021 <- factor( ifelse( d_2002$F21F91 == 1, "Heterosexual",
                                                  ifelse( d_2002$F21F91 == 2, "Homosexual",
                                                          ifelse( d_2002$F21F91 == 3, "Bisexual", "Other" ) ) ),
                                          levels = c( "Heterosexual", "Homosexual", "Bisexual", "Other" ) )
table( d_2002$sexual_identity_2021, useNA = "always" )

# sex
table( d_2002$kon, useNA = "always" )
d_2002$sex <- factor( ifelse( d_2002$kon == 1, "Male", "Female" ),
                      levels = c( "Male", "Female" ) )
table( d_2002$sex, useNA = "always" )

# age
summary( d_2002$F2alder )
d_2002$age_2002 <- d_2002$F2alder
d_2002$age_2010 <- d_2002$age_2002 + 8
```

#### 2.2. SPHC 2006
```{r}
d_2006 <- read_sas("/Volumes/LGBT Project data/sphc06_10_14_21.sas7bdat")

# sexual identity in 2010
table( d_2006$F10F103, useNA = "always" )
d_2006$F10F103[ d_2006$F10F103 == 9 ] <- NA
d_2006$sexual_identity_2010 <- factor( ifelse( d_2006$F10F103 == 1, "Heterosexual",
                                                  ifelse( d_2006$F10F103 == 2, "Homosexual",
                                                          ifelse( d_2006$F10F103 == 3, "Bisexual", "Other" ) ) ),
                                          levels = c( "Heterosexual", "Homosexual", "Bisexual", "Other" ) )
table( d_2006$sexual_identity_2010, useNA = "always" )

# sexual identity in 2014
table( d_2006$F14F103, useNA = "always" )
d_2006$sexual_identity_2014 <- factor( ifelse( d_2006$F14F103 == 1, "Heterosexual",
                                                  ifelse( d_2006$F14F103 == 2, "Homosexual",
                                                          ifelse( d_2006$F14F103 == 3, "Bisexual", "Other" ) ) ),
                                          levels = c( "Heterosexual", "Homosexual", "Bisexual", "Other" ) )
table( d_2006$sexual_identity_2014, useNA = "always" )

# sexual identity in 2021
table( d_2006$F21F91, useNA = "always" )
d_2006$sexual_identity_2021 <- factor( ifelse( d_2006$F21F91 == 1, "Heterosexual",
                                                  ifelse( d_2006$F21F91 == 2, "Homosexual",
                                                          ifelse( d_2006$F21F91 == 3, "Bisexual", "Other" ) ) ),
                                          levels = c( "Heterosexual", "Homosexual", "Bisexual", "Other" ) )
table( d_2006$sexual_identity_2021, useNA = "always" )

# sex
table( d_2006$kon, useNA = "always" )
d_2006$sex <- factor( ifelse( d_2006$kon == 1, "Male", "Female" ),
                      levels = c( "Male", "Female" ) )
table( d_2006$sex, useNA = "always" )

# age
summary( d_2006$F6alder )
d_2006$age_2006 <- d_2006$F6alder
d_2006$age_2010 <- d_2006$age_2006 + 4
```

#### 2.3. SPHC 2010
```{r}
load("/Volumes/LGBT Project data/d_2010.RData")

# sexual identity in 2010
table( d_2010$F10U87G78, useNA = "always" )
d_2010$F10U87G78[ d_2010$F10U87G78 == 9 ] <- NA
d_2010$sexual_identity_2010 <- factor( ifelse( d_2010$F10U87G78 == 1, "Heterosexual",
                                                  ifelse( d_2010$F10U87G78 == 2, "Homosexual",
                                                          ifelse( d_2010$F10U87G78 == 3, "Bisexual", "Other" ) ) ),
                                          levels = c( "Heterosexual", "Homosexual", "Bisexual", "Other" ) )
table( d_2010$sexual_identity_2010, useNA = "always" )

# sexual identity in 2014
table( d_2010$F14F103, useNA = "always" )
d_2010$sexual_identity_2014 <- factor( ifelse( d_2010$F14F103 == 1, "Heterosexual",
                                                  ifelse( d_2010$F14F103 == 2, "Homosexual",
                                                          ifelse( d_2010$F14F103 == 3, "Bisexual", "Other" ) ) ),
                                          levels = c( "Heterosexual", "Homosexual", "Bisexual", "Other" ) )
table( d_2010$sexual_identity_2014, useNA = "always" )

# sexual identity in 2021
table( d_2010$F21F91, useNA = "always" )
d_2010$sexual_identity_2021 <- factor( ifelse( d_2010$F21F91 == 1, "Heterosexual",
                                                  ifelse( d_2010$F21F91 == 2, "Homosexual",
                                                          ifelse( d_2010$F21F91 == 3, "Bisexual", "Other" ) ) ),
                                          levels = c( "Heterosexual", "Homosexual", "Bisexual", "Other" ) )
table( d_2010$sexual_identity_2021, useNA = "always" )

# sex
table( d_2010$kon, useNA = "always" )
d_2010$sex <- factor( ifelse( d_2010$kon == 1, "Male", "Female" ),
                      levels = c( "Male", "Female" ) )
table( d_2010$sex, useNA = "always" )

# age
summary( d_2010$F10alder )
d_2010$age_2010 <- d_2010$F10alder
```

#### 2.4. Merge data across surveys
```{r}
# there are 999 individuals who participated in both SPHC 2002 and 2006
# their follow-up data were included only once

variable_list <- c( "lopnr", "sexual_identity_2010", "sexual_identity_2014", "sexual_identity_2021", "age_2010", "sex" )
  
m1 <- d_2002[ , variable_list ] %>%
  mutate( year = rep( "SPHC 2002" ) )

m2 <- d_2006[ , variable_list ] %>%
  mutate( year = rep( "SPHC 2006" ) )

m12 <- rbind( m1, m2 ) %>%
  group_by( lopnr ) %>%
  filter( !( year == "SPHC 2002" & any( year == "SPHC 2006" ) ) ) %>%
  ungroup()

# age
summary( m12$age_2010 )
m12$age_exp <- factor( ifelse( m12$age_2010 <= 29, "<=29",
                               ifelse( m12$age_2010 >=30 & m12$age_2010 <= 44, "30-44",
                                       ifelse( m12$age_2010 >= 45 & m12$age_2010 <= 59, "45-59", ">=60" ) ) ),
                       levels = c( "45-59", "<=29", "30-44", ">=60" ) )
table( m12$age_exp, useNA = "always" )

# sex
summary( m12$sex )
m12$sex_exp <-relevel( m12$sex, ref = "Male" )

# year
m12$year <- as.factor( m12$year )
table( m12$year, useNA = "always" )

summary( m12 )
d_2002_2006_follow_up <- m12 %>%
  select( "sexual_identity_2010", "sexual_identity_2014", "sexual_identity_2021", "age_exp", "sex_exp", "year" )


##### obtain follow-up data in 2010-2021 for SPHC 2010 #####

d_2010 <- d_2010 %>% mutate( year = rep( "SPHC 2010" ) )

# age
summary( d_2010$age_2010 )
d_2010$age_exp <- factor( ifelse( d_2010$age_2010 <= 29, "<=29", 
                              ifelse( d_2010$age_2010 >=30 & d_2010$age_2010 <= 44, "30-44",
                                      ifelse( d_2010$age_2010 >= 45 & d_2010$age_2010 <= 59, "45-59", ">=60" ) ) ),
                      levels = c( "45-59", "<=29", "30-44", ">=60" ) )
table( d_2010$age_exp, useNA = "always" )

# sex
summary( d_2010$sex )
d_2010$sex_exp <- relevel( d_2010$sex, ref = "Male" )

summary( d_2010 )
d_2010_follow_up <- d_2010 %>%
  select( "sexual_identity_2010", "sexual_identity_2014", "sexual_identity_2021", "age_exp", "sex_exp", "year" )


##### merge follow-up data in 2010-2021 for SPHC 2002, 2006, and 2010 #####

d_2002_2006_2010_follow_up <- rbind( d_2002_2006_follow_up, d_2010_follow_up )

# define outcome
d_2002_2006_2010_follow_up$sexual_identity_fluidity <- ifelse(
    d_2002_2006_2010_follow_up$sexual_identity_2010 != d_2002_2006_2010_follow_up$sexual_identity_2014 | 
    d_2002_2006_2010_follow_up$sexual_identity_2010 != d_2002_2006_2010_follow_up$sexual_identity_2021 |
    d_2002_2006_2010_follow_up$sexual_identity_2014 != d_2002_2006_2010_follow_up$sexual_identity_2021, 
    1, 0 ) # any change of sexual identity in 2010-2021

d_sexual_identity_fluidity <- d_2002_2006_2010_follow_up[ 
  !is.na( d_2002_2006_2010_follow_up$sexual_identity_2010 ) &
    !is.na( d_2002_2006_2010_follow_up$sexual_identity_2014 ) &
    !is.na( d_2002_2006_2010_follow_up$sexual_identity_2021 ), c( "sexual_identity_2010", "sexual_identity_2014", "sexual_identity_2021", "age_exp", "sex_exp", "year" ) ]
nrow( d_sexual_identity_fluidity ) # 34,815 provided data on sexual identity in 2010-2021
n_miss( d_sexual_identity_fluidity ) # 0 missing
```

### 3. Make Fluidity Plots
#### 3.1. Prepare dataset for plotting
##### 3.1.1. Among all participants
```{r}
# change in sexual identity from 2010 to 2021
m123_all <- d_sexual_identity_fluidity %>%
  count( sexual_identity_2010, sexual_identity_2014, sexual_identity_2021, name = "number" ) %>%
  mutate( percent = number / sum( number ) )

m123_all <- m123_all %>%
  group_by( sexual_identity_2010 ) %>%
  mutate( percent_for_label = number / sum( number ) ) %>%
  ungroup()

m123_all$label <- paste0( prettyNum( m123_all$number, big.mark = "," , preserve.width = "none" ), 
                          " (", sprintf( "%.1f", m123_all$percent_for_label*100 ), "%)" 
                          )

for ( col in 1:3 ) {
  m123_all[[col]] <- factor( m123_all[[col]], levels = c( "Homosexual", "Bisexual", "Other", "Heterosexual" ) ) 
}

m123_all_long <- to_lodes_form( m123_all, axes = 1:3, key = "year", value = "sexual_identity", id = "Cohort", diffuse = sexual_identity_2010 ) # convert to long format
```

##### 3.1.2. By age
```{r}
summary( d_sexual_identity_fluidity$age_exp )

m123_age <- d_sexual_identity_fluidity %>%
  count( sexual_identity_2010, sexual_identity_2014, sexual_identity_2021, age_exp ) %>%
  rename( number = n ) %>%
  group_by( age_exp ) %>%
  mutate( percent = number / sum( number ) ) %>%
  ungroup()

for ( col in 1:3 ) {
  m123_age[[col]] <- factor( m123_age[[col]], levels = c( "Homosexual", "Bisexual", "Other", "Heterosexual" ) ) 
}


# 18-29 yrs
m123_age_18 <- m123_age %>%
  filter( age_exp == "<=29" ) %>%
  group_by( sexual_identity_2010 ) %>%
  mutate( percent_for_label = number / sum( number ) ) %>%
  ungroup()

m123_age_18$label <- paste0( prettyNum( m123_age_18$number, big.mark = "," , preserve.width = "none" ), " (", 
                          sprintf( "%.1f", m123_age_18$percent_for_label*100 ),
                          "%)" )

m123_age_18_long <- to_lodes_form( m123_age_18, axes = 1:3, key = "year", value = "sexual_identity", id = "Cohort", diffuse = sexual_identity_2010 )


# 30-44 yrs
m123_age_30 <- m123_age %>%
  filter( age_exp == "30-44" ) %>%
  group_by( sexual_identity_2010 ) %>%
  mutate( percent_for_label = number / sum( number ) ) %>%
  ungroup()

m123_age_30$label <- paste0( prettyNum( m123_age_30$number, big.mark = "," , preserve.width = "none" ), " (", 
                          sprintf( "%.1f", m123_age_30$percent_for_label*100 ),
                          "%)" )

m123_age_30_long <- to_lodes_form( m123_age_30, axes = 1:3, key = "year", value = "sexual_identity", id = "Cohort", diffuse = sexual_identity_2010 )


# 45-59 yrs
m123_age_45 <- m123_age %>%
  filter( age_exp == "45-59" ) %>%
  group_by( sexual_identity_2010 ) %>%
  mutate( percent_for_label = number / sum( number ) ) %>%
  ungroup()

m123_age_45$label <- paste0( prettyNum( m123_age_45$number, big.mark = "," , preserve.width = "none" ), " (", 
                          sprintf( "%.1f", m123_age_45$percent_for_label*100 ),
                          "%)" )

m123_age_45_long <- to_lodes_form( m123_age_45, axes = 1:3, key = "year", value = "sexual_identity", id = "Cohort", diffuse = sexual_identity_2010 )


# >=60 yrs
m123_age_60 <- m123_age %>%
  filter( age_exp == ">=60" ) %>%
  group_by( sexual_identity_2010 ) %>%
  mutate( percent_for_label = number / sum( number ) ) %>%
  ungroup()

m123_age_60$label <- paste0( prettyNum( m123_age_60$number, big.mark = "," , preserve.width = "none" ), " (", 
                          sprintf( "%.1f", m123_age_60$percent_for_label*100 ),
                          "%)" )

m123_age_60_long <- to_lodes_form( m123_age_60, axes = 1:3, key = "year", value = "sexual_identity", id = "Cohort", diffuse = sexual_identity_2010 )
```

#### 3.2. Plotting
##### 3.2.1. Among all participants
```{r}
# version 1
ggplot( m123_all_long,
        aes( x = year,
             stratum = sexual_identity, 
             alluvium = Cohort,
             y = percent ) ) +
  geom_alluvium( aes( fill = sexual_identity_2010 ), curve_type = "cubic" ) +
  geom_stratum( aes( fill = sexual_identity_2010 ) ) +
  geom_text( stat = "stratum", aes( label = after_stat( stratum ) ), family = "Arial", size = 12 / .pt ) +
  scale_fill_manual( values = c( "#C77CFF", "#00BFC4", "#7CAE00", "#F8766D" ),
                     breaks = c( "Heterosexual", "Homosexual", "Bisexual", "Other" ),
                     na.value = NA ) +
  scale_x_discrete( limits = c( "sexual_identity_2010", "sexual_identity_2014", "sexual_identity_2021" ), labels = c( "2010", "2014", "2021" ), expand = c( 0, 0 ) ) +
  theme_classic() +
  scale_y_break( c( 0.12, 0.9 ), scales = 10, space = 0.3 ) +
  scale_y_continuous( labels = scales::percent, breaks = c( 0, 0.1, 0.9, 0.95, 1 ) ) +
  labs( x = NULL, y = NULL) +
  theme( axis.text.x = element_text( family = "Arial", size = 11 ),
         axis.text.y = element_text( family = "Arial", size = 11 ),
         legend.title = element_blank(),
         legend.text = element_text( family = "Arial", size = 10 ), 
         legend.key = element_rect( colour = "white" ),
         legend.position = "bottom",
         axis.ticks.x.bottom = element_blank(),
         axis.line.x.bottom = element_blank(),
         axis.text.y.right = element_blank(),
         axis.ticks.y.right = element_blank(),
         axis.line.y.right = element_blank()
         ) +
  geom_text_repel( stat = "alluvium",
                   aes( x = as.integer( year ) + 0.15,
                        color = factor( sexual_identity_2010, 
                                        levels = c( "Other", "Bisexual", "Homosexual", "Heterosexual" ) ),
                        label = ifelse( year == "sexual_identity_2021" & 
                                          ( percent_for_label >= 0.1 | number >= 50 ), label, NA ) ),
                   size = 4.5,
                   family = "Arial",
                   segment.linetype = 3,
                   segment.size = 0.55,
                   direction = "y",
                   nudge_x = 0.45,
                   box.padding = 0.5,
                   point.padding = 0,
                   show.legend = FALSE,
                   seed = 123
                   )
```

```{r}
# version 2
ggplot( m123_all_long,
        aes( x = year,
             stratum = sexual_identity, 
             alluvium = Cohort,
             y = percent ) ) +
  geom_alluvium( aes( fill = sexual_identity_2010 ), curve_type = "cubic" ) +
  geom_stratum( aes( fill = sexual_identity_2010 ) ) +
  geom_text( stat = "stratum", aes( label = after_stat( stratum ) ), family = "Arial", size = 12 / .pt ) +
  scale_fill_manual( values = c( "#C77CFF", "#00BFC4", "#7CAE00", "#F8766D" ),
                     breaks = c( "Heterosexual", "Homosexual", "Bisexual", "Other" ),
                     na.value = NA ) +
  scale_x_discrete( limits = c( "sexual_identity_2010", "sexual_identity_2014", "sexual_identity_2021" ), labels = c( "2010", "2014", "2021" ), expand = c( 0, 0 ) ) +
  expand_limits( x = 3.2 ) +
  theme_classic() +
  scale_y_continuous( labels = scales::percent, breaks = seq( 0, 1, by = 0.1 ) ) +
  labs( x = NULL, y = NULL) +
  theme( axis.text.x = element_text( family = "Arial", size = 11 ),
         axis.text.y = element_text( family = "Arial", size = 11 ),
         legend.title = element_blank(),
         legend.position = "none",
         axis.ticks.x.bottom = element_blank(),
         axis.line.x.bottom = element_blank(),
         axis.text.y.right = element_blank(),
         axis.ticks.y.right = element_blank(),
         axis.line.y.right = element_blank()
  )
```

##### 3.2.2. By age
```{r}
# 18-29 years 
ggplot( m123_age_18_long,
        aes( x = year,
             stratum = sexual_identity, 
             alluvium = Cohort,
             y = percent ) ) +
  geom_alluvium( aes( fill = sexual_identity_2010 ), curve_type = "cubic" ) +
  geom_stratum( aes( fill = sexual_identity_2010 ) ) +
  geom_text( stat = "stratum", aes( label = after_stat( stratum ) ), family = "Arial", size = 12 / .pt ) +
  scale_fill_manual( values = c( "#C77CFF", "#00BFC4", "#7CAE00", "#F8766D" ),
                     breaks = c( "Heterosexual", "Homosexual", "Bisexual", "Other" ),
                     na.value = NA ) +
  scale_x_discrete( limits = c( "sexual_identity_2010", "sexual_identity_2014", "sexual_identity_2021" ), labels = c( "2010", "2014", "2021" ), expand = c( 0, 0 ) ) +
  theme_classic() +
  scale_y_break( c( 0.12, 0.86 ), scales = 10, space = 0.3 ) +
  scale_y_continuous( labels = scales::percent, breaks = c( 0, 0.1, 0.9, 0.95, 1 ) ) +
  labs( x = NULL, y = NULL) +
  theme( axis.text.x = element_text( family = "Arial", size = 11 ),
         axis.text.y = element_text( family = "Arial", size = 11 ),
         legend.title = element_blank(),
         legend.text = element_text( family = "Arial", size = 10 ), 
         legend.key = element_rect( colour = "white" ),
         legend.position = "bottom",
         axis.ticks.x.bottom = element_blank(),
         axis.line.x.bottom = element_blank(),
         axis.text.y.right = element_blank(),
         axis.ticks.y.right = element_blank(),
         axis.line.y.right = element_blank()
         ) +
  geom_text_repel( stat = "alluvium",
                   aes( x = as.integer( year ) + 0.15,
                        color = factor( sexual_identity_2010, levels = c( "Other", "Bisexual", "Homosexual", "Heterosexual" ) ),
                        label = ifelse( year == "sexual_identity_2021" & ( percent_for_label >= 0.1 | number >= 6 ), label, NA ) ),
                   size = 4.5,
                   family = "Arial",
                   segment.linetype = 3,
                   segment.size = 0.55,
                   direction = "y",
                   nudge_x = 0.4,
                   box.padding = 0.5,
                   point.padding = 0,
                   show.legend = FALSE,
                   seed = 123
                   )


# 30-44 years
ggplot( m123_age_30_long,
        aes( x = year,
             stratum = sexual_identity, 
             alluvium = Cohort,
             y = percent ) ) +
  geom_alluvium( aes( fill = sexual_identity_2010 ), curve_type = "cubic" ) +
  geom_stratum( aes( fill = sexual_identity_2010 ) ) +
  geom_text( stat = "stratum", aes( label = after_stat( stratum ) ), family = "Arial", size = 12 / .pt ) +
  scale_fill_manual( values = c( "#C77CFF", "#00BFC4", "#7CAE00", "#F8766D" ),
                     breaks = c( "Heterosexual", "Homosexual", "Bisexual", "Other" ),
                     na.value = NA ) +
  scale_x_discrete( limits = c( "sexual_identity_2010", "sexual_identity_2014", "sexual_identity_2021" ), labels = c( "2010", "2014", "2021" ), expand = c( 0, 0 ) ) +
  theme_classic() +
  scale_y_break( c( 0.12, 0.915 ), scales = 10, space = 0.3 ) +
  scale_y_continuous( labels = scales::percent, breaks = c( 0, 0.1, 0.9, 0.95, 1 ) ) +
  labs( x = NULL, y = NULL) +
  theme( axis.text.x = element_text( family = "Arial", size = 11 ),
         axis.text.y = element_text( family = "Arial", size = 11 ),
         legend.title = element_blank(),
         legend.text = element_text( family = "Arial", size = 10 ), 
         legend.key = element_rect( colour = "white" ),
         legend.position = "bottom",
         axis.ticks.x.bottom = element_blank(),
         axis.line.x.bottom = element_blank(),
         axis.text.y.right = element_blank(),
         axis.ticks.y.right = element_blank(),
         axis.line.y.right = element_blank()
         ) +
  geom_text_repel( stat = "alluvium",
                   aes( x = as.integer( year ) + 0.15,
                        color = factor( sexual_identity_2010, levels = c( "Other", "Bisexual", "Homosexual", "Heterosexual" ) ),
                        label = ifelse( year == "sexual_identity_2021" & ( percent_for_label >= 0.1 | number >= 12 ), label, NA ) ),
                   size = 4.5,
                   family = "Arial",
                   segment.linetype = 3,
                   segment.size = 0.55,
                   direction = "y",
                   nudge_x = 0.4,
                   box.padding = 0.5,
                   point.padding = 0,
                   show.legend = FALSE,
                   seed = 123
                   )


# 45-59 years
ggplot( m123_age_45_long,
        aes( x = year,
             stratum = sexual_identity, 
             alluvium = Cohort,
             y = percent ) ) +
  geom_alluvium( aes( fill = sexual_identity_2010 ), curve_type = "cubic" ) +
  geom_stratum( aes( fill = sexual_identity_2010 ) ) +
  geom_text( stat = "stratum", aes( label = after_stat( stratum ) ), family = "Arial", size = 12 / .pt ) +
  scale_fill_manual( values = c( "#C77CFF", "#00BFC4", "#7CAE00", "#F8766D" ),
                     breaks = c( "Heterosexual", "Homosexual", "Bisexual", "Other" ),
                     na.value = NA ) +
  scale_x_discrete( limits = c( "sexual_identity_2010", "sexual_identity_2014", "sexual_identity_2021" ), labels = c( "2010", "2014", "2021" ), expand = c( 0, 0 ) ) +
  theme_classic() +
  scale_y_break( c( 0.12, 0.916 ), scales = 10, space = 0.3 ) +
  scale_y_continuous( labels = scales::percent, breaks = c( 0, 0.1, 0.9, 0.95, 1 ) ) +
  labs( x = NULL, y = NULL) +
  theme( axis.text.x = element_text( family = "Arial", size = 11 ),
         axis.text.y = element_text( family = "Arial", size = 11 ),
         legend.title = element_blank(),
         legend.text = element_text( family = "Arial", size = 10 ), 
         legend.key = element_rect( colour = "white" ),
         legend.position = "bottom",
         axis.ticks.x.bottom = element_blank(),
         axis.line.x.bottom = element_blank(),
         axis.text.y.right = element_blank(),
         axis.ticks.y.right = element_blank(),
         axis.line.y.right = element_blank()
         ) +
  geom_text_repel( stat = "alluvium",
                   aes( x = as.integer( year ) + 0.15,
                        color = factor( sexual_identity_2010, levels = c( "Other", "Bisexual", "Homosexual", "Heterosexual" ) ),
                        label = ifelse( year == "sexual_identity_2021" & ( percent_for_label >= 0.1 | number >= 15 ), label, NA ) ),
                   size = 4.5,
                   family = "Arial",
                   segment.linetype = 3,
                   segment.size = 0.55,
                   direction = "y",
                   nudge_x = 0.4,
                   box.padding = 0.5,
                   point.padding = 0,
                   show.legend = FALSE,
                   seed = 123
                   )


# >=60
ggplot( m123_age_60_long,
        aes( x = year,
             stratum = sexual_identity, 
             alluvium = Cohort,
             y = percent ) ) +
  geom_alluvium( aes( fill = sexual_identity_2010 ), curve_type = "cubic" ) +
  geom_stratum( aes( fill = sexual_identity_2010 ) ) +
  geom_text( stat = "stratum", aes( label = after_stat( stratum ) ), family = "Arial", size = 12 / .pt ) +
  scale_fill_manual( values = c( "#C77CFF", "#00BFC4", "#7CAE00", "#F8766D" ),
                     breaks = c( "Heterosexual", "Homosexual", "Bisexual", "Other" ),
                     na.value = NA ) +
  scale_x_discrete( limits = c( "sexual_identity_2010", "sexual_identity_2014", "sexual_identity_2021" ), labels = c( "2010", "2014", "2021" ), expand = c( 0, 0 ) ) +
  theme_classic() +
  scale_y_break( c( 0.12, 0.881 ), scales = 10, space = 0.3 ) +
  scale_y_continuous( labels = scales::percent, breaks = c( 0, 0.1, 0.9, 0.95, 1 ) ) +
  labs( x = NULL, y = NULL) +
  theme( axis.text.x = element_text( family = "Arial", size = 11 ),
         axis.text.y = element_text( family = "Arial", size = 11 ),
         legend.title = element_blank(),
         legend.text = element_text( family = "Arial", size = 10 ), 
         legend.key = element_rect( colour = "white" ),
         legend.position = "bottom",
         axis.ticks.x.bottom = element_blank(),
         axis.line.x.bottom = element_blank(),
         axis.text.y.right = element_blank(),
         axis.ticks.y.right = element_blank(),
         axis.line.y.right = element_blank()
         ) +
  geom_text_repel( stat = "alluvium",
                   aes( x = as.integer( year ) + 0.15,
                        color = factor( sexual_identity_2010, levels = c( "Other", "Bisexual", "Homosexual", "Heterosexual" ) ),
                        label = ifelse( year == "sexual_identity_2021" & ( percent_for_label >= 0.1 | number >= 15 ), label, NA ) ),
                   size = 4.5,
                   family = "Arial",
                   segment.linetype = 3,
                   segment.size = 0.55,
                   direction = "y",
                   nudge_x = 0.4,
                   box.padding = 0.5,
                   point.padding = 0,
                   show.legend = FALSE,
                   seed = 123
                   )
```

```{r}
# check status
renv::status()

# record R package versions
renv::snapshot()
```